Towards Proof-Reconstruction of Problems in Classical Propositional Logic in Agda

### 2017-07-28 paper/main.tex
minor changes
minor changes.

\section{Introduction}
Proof reconstruction is a hard labor, it integrates two
complex systems. On one hand, we have the automatic theorem provers (henceforth
ATPs) that depending on the specification logic and background theories exist
many alternatives. They are SAT solvers (e.g. \verb!zChaff!~\cite{Moskewicz2001} and \verb!MiniSat!~\cite{Een2004}) to prove unsatisfiability of CNF formulas,
QBF solvers (e.g.~\verb!GhostQ!~\cite{Klieber2014} and ~\verb!DepQBF!~\cite{Lonsing2017}) to prove satisfiability and invalidity of quantified Boolean
formulas, SMT solvers (e.g \verb!CVC4!~\cite{Barrett2011}, \verb!veriT!~\cite{bouton2009}, \verb!Z3!~\cite{DeMoura2008}) to
prove unsatisfiability of formulas from first-order logic with theories,  and
finally, the ATPs to prove validity of formulas from first-order logic with
equality (e.g.~\verb!E!~\cite{Schulz:AICOM-2002},~\verb!Gandalf!~\cite{Tammet1997},~\verb!leanCoP!~\cite{Otten2008},~\verb!Metis!~\cite{hurd2003first},~\verb!SPASS!~\cite{Weidenbach2009} and ~\verb!Vampire!~\cite{Riazanov1999}),
high-order logic (e.g.~\verb!Leo-II!,~\verb!Satallax!), intuitonistic logic
(e.g.~\verb!ileanCoP!~\cite{Otten2008},~\verb!JProver!~\cite{Schmitt2001},~\verb!Gandalf!~\cite{Tammet1997}), among others.

For proof reconstruction, we are interested in ATP tools that can deliver proof
objects in a consistent format to work with, that is, a full script describing
step-by-step the table of the derivation to get the actual proof. Unfortunately,
just a few of the ATPs generate in their outputs a proof object.

For problems in classical propositional logic (henceforth CPL), from a list of
at least forty\footnote{ATPs available in the website  \texttt{SystemOnTPTP}
  from the TPTP World} ATPs, just a few deliver proofs (e.g.
\verb!CVC4!~\cite{Barrett2011}, \verb!SPASS!~\cite{Weidenbach2009},
\verb!Waldmeister!~\cite{hillenbrand1997}) and a little bit less proofs in file
format like TSTP (e.g. \verb!E!~\cite{Schulz:AICOM-2002}, \verb!Metis!~\cite{hurd2003first}, \verb!Vampire!~\cite{Riazanov1999}, \verb!Z3!).

On the other hand, we have the proof checkers, Interactive Theorem Provers
%(henceforth ITP)
or proof assistants (e.g. \verb!Agda!~\cite{agdateam}, \verb!Coq!~\cite{coqteam}, \verb!Isabelle!~\cite{paulson1994isabelle}, \verb!HOL4!~\cite{norrish2007hol}). These programs allow us to define the formal language for the problems like operators, logic variables, axioms, and theorems. They can assist us to check and validate the proof script delivered by the provers. Because the formalism in the source (the proof generated by the ATP) is not necessarily the same formalism in the target (proof reconstructed in the proof assistant), then the reconstruction of a proof involves a deep understanding around the deduction process done by the ATP, for translating that deduction later into the proof assistant.

>>>>

\section{Introduction}
Proof reconstruction is a hard labor, it integrates two
complex systems. On one hand, we have the automatic theorem provers (henceforth
ATPs) that depending on the specification logic and background theories exist
many alternatives. They are SAT solvers (e.g. \verb!zChaff!~\cite{Moskewicz2001} and \verb!MiniSat!~\cite{Een2004}) to prove unsatisfiability of CNF formulas,
QBF solvers (e.g. \verb!GhostQ!~\cite{Klieber2014} and~\verb!DepQBF!~\cite{Lonsing2017}) to prove satisfiability and invalidity of quantified Boolean
formulas, SMT solvers (e.g \verb!CVC4!~\cite{Barrett2011}, \verb!veriT!~\cite{bouton2009}, and \verb!Z3!~\cite{DeMoura2008}) to
prove unsatisfiability of formulas from first-order logic with theories,  and
finally, the ATPs to prove validity of formulas from first-order logic with
equality (e.g. \verb!E!~\cite{Schulz:AICOM-2002},~\verb!Gandalf!~\cite{Tammet1997},~\verb!leanCoP!~\cite{Otten2008},~\verb!Metis!~\cite{hurd2003first},~\verb!SPASS!~\cite{Weidenbach2009} and ~\verb!Vampire!~\cite{Riazanov1999}),
high-order logic (e.g. \verb!Leo-II!~\cite{Benzmuller2008} and \verb!Satallax!~\cite{Brown2012}), intuitonistic logic
(e.g. \verb!ileanCoP!~\cite{Otten2008},~\verb!JProver!~\cite{Schmitt2001}, and \verb!Gandalf!~\cite{Tammet1997}), among others.

For proof reconstruction, we need from the ATP tools proof objects in a consistent format to work with, that is, a full script describing
step-by-step the table of the derivation to get the actual proof. Unfortunately,
not all ATPs can deliver proofs or some useful output for reconstruct a proof.
For problems in classical propositional logic (henceforth CPL), from a list of
at least forty\footnote{ATPs available in the website  \texttt{SystemOnTPTP} from the TPTP World} ATPs, just a few were able to deliver proofs (e.g.
\verb!CVC4!~\cite{Barrett2011}, \verb!SPASS!, and
\verb!Waldmeister!~\cite{hillenbrand1997}) and a little bit less generate proofs in a file
format like TSTP~\cite{sutcliffe2004tstp} (e.g. \verb!E!~\cite{Schulz:AICOM-2002}, \verb!Metis!~\cite{hurd2003first}, \verb!Vampire!, and \verb!Z3!) or LFSC~\cite{Stump2008}.

On the other hand, we have the proof checkers, Interactive Theorem Provers
%(henceforth ITP)
or proof assistants (e.g. \verb!Agda!~\cite{agdateam}, \verb!Coq!~\cite{coqteam}, \verb!Isabelle!~\cite{paulson1994isabelle}, \verb!HOL4!~\cite{norrish2007hol}). These programs allow us to define the formal language for the problems like operators, logic variables, axioms, and theorems. They can assist us to check and validate the proof script delivered by the provers. Because the formalism in the source (the proof generated by the ATP) is not necessarily the same formalism in the target (proof reconstructed in the proof assistant), the reconstruction of a proof involves a deep understanding around the deduction process done by the ATP, for translating that deduction later into the proof assistant, the so called ``reverse engineering''.
